---
title: "Queue TP Reduction"
author: "Logan Wu"
date: "`r format(lubridate::today(), '%B %e, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(patchwork)
library(readxl)
library(lubridate)
library(distributional)
library(future.apply)
library(truncdist)
library(scales)
plan(multisession)
z = 1.96

for(x in list.files("R", full.names=T)) {source(x)}
```

This RMarkdown illustrates the effectiveness of queueing strategy on transmission potential (TP) reduction in a contact tracing context.

```{r}
data_dir = "data"
linelist_file = list.files(data_dir, "Linelist_Cases")

if (length(linelist_file) == 1) {
  linelist = suppressWarnings(read_excel(file.path(data_dir, linelist_file))) %>%
    mutate_if(is.POSIXct, as_date)
  # Interview delay plot
  weekly_delays = linelist %>%
    mutate(total_delay = InterviewDate - DiagnosisDate,
           total_delay = ifelse(total_delay < 0, 0, total_delay),
           total_delay = ifelse(total_delay > 7, 7, total_delay),
           total_delay = total_delay %>% as_factor() %>% fct_recode("<1"="0", "7+"="7"),
           DiagnosisDate = as_date(round(as.numeric(DiagnosisDate) / 7) * 7)) %>%
    filter(DiagnosisDate > as_date("2020-01-01")) %>%
    count(DiagnosisDate, total_delay) %>%
    group_by(DiagnosisDate) %>%
    mutate(pct = n / sum(n)) %>%
    ungroup()
  weekly_counts = linelist %>%
    mutate(DiagnosisDate = as_date(round(as.numeric(DiagnosisDate) / 7) * 7)) %>%
    filter(DiagnosisDate > as_date("2020-01-01")) %>%
    count(DiagnosisDate) %>%
    mutate(scaled_n = n / max(n))
  
  ggplot() +
    geom_col(data=weekly_delays, aes(x = DiagnosisDate, y = pct, fill=total_delay)) +
    geom_line(data=weekly_counts, aes(x=DiagnosisDate, y=scaled_n), color="white", size=2) +
    geom_line(data=weekly_counts, aes(x=DiagnosisDate, y=scaled_n), color="black", size=0.5) +
    scale_fill_brewer(direction=-1, palette="Spectral", na.value="grey") +
    theme_minimal() +
    theme(legend.position="bottom") +
    labs(title = "Victorian COVID-19 interview delays",
         x = "Diagnosis week",
         y = "Percent of interviewed cases",
         fill = "Interview delay (days)\nafter Dept. notification") +
    scale_x_date(date_breaks = "3 months", labels = function(x) format(x, "%b '%y")) +
    scale_y_continuous(labels = scales::percent,
                       sec.axis = sec_axis( trans=~.*max(weekly_counts$n), name="Weekly cases"))
  ggsave("plot6.svg", width=150, height=75, units="mm")
}
```


Read in the empirical notification delay

```{r}
# notification_delays = read_csv("notification_delays.csv")
test_turnaround_time = read_csv("data-public/test_turnaround_time.csv")
```

Define our priority functions.

```{r}
#' Oldest swab priority function
#'
#' Arranges the current queue by the order they should be interviewed in. Should
#' always begin with the interview eligibility criteria so interviewable cases
#' are at the top.
#'
#' @param x Data frame of the current queue state
oldest_swab = function(x) {
  x %>%
    arrange(
      # Whether case is eligible to be interviewed
      desc(eligible_for_interview),
      # Priorities, in order of appearance
      swab_date # oldest first
    )
}

newest_swab = function(x) {
  x %>%
    arrange(
      # Whether case is eligible to be interviewed
      desc(eligible_for_interview),
      # Priorities, in order of appearance
      desc(swab_date) # newest first
    )
}

newest_notification = function(x) {
  x %>%
    arrange(
      # Whether case is eligible to be interviewed
      desc(eligible_for_interview),
      # Priorities, in order of appearance
      desc(notification_date)#, # newest first
      # vaccinated # vaccinated FALSE first
    )
}

random_swab <- function(x) {
  x %>%
    mutate(ix = runif(n())) %>%
    arrange(
      # Whether case is eligible to be interviewed
      desc(eligible_for_interview),
      # Priorities, in order of appearance
      ix
    ) %>%
    select(-ix)
}

new_swab_vaccine <- function(x, sim_day, notification_time) {
  x %>%
    arrange(
      # Whether case is eligible to be interviewed
      desc(eligible_for_interview),
      # Priorities, in order of appearance
      desc(swab_date), # newest first
      vaccinated # vaccinated FALSE first
    )
}
```

Create a dataframe representing the linelist and process the queue day by day.

```{r}
#' Run queue simulation
#' 
#' Function to run the simulation. Written as a function to make multiple runs
#' more convenient.
#'
#' @param capacity_ratio Daily capacity is this fraction of the mean incoming case rate.
#' @param max_interview_delay Cases cannot be interviewed greater than this number of days after their notification.
#' @param prop_priority Proportion of cases that can be assigned to a 'priority' group. Currently has no effect on TP.
#' @param prop_time_delay Proportion of cases that are notified after hours and cannot be called the same day.
#' @param f_priority Prioritisation function to be applied before deciding which cases to interview on a given day.
#' @param proportion_cases_vaccinated Simular to prop_priority, proportion of cases that should be assigned a 'vaccinated' flag before prioritisation.
#' @param days_samples Approximate number of days to sample arrivals from. The actual number can vary due to arrival sampling.
#' @param days_burnin Approximate number of days to sample and discard to initialise the queue.
#' @param days_burnout Days to simulate after the sample days so cases near the end can be interviewed.
#' @param mean_rate Mean arrival rate.
sim_tracing = function(capacity_ratio = 0.8,
                       max_interview_delay = 5,
                       prop_priority = 0.4,
                       prop_time_delay = 0.2,
                       f_priority = f_priority,
                       proportion_cases_vaccinated = 0.45,
                       days_samples = 100,
                       days_burnin = 10,
                       days_burnout = 5,
                       mean_rate = 20) {
  # print("Executing...")
  n_samples = days_samples * mean_rate
  n_burnin = days_burnin * mean_rate
  n_burnout = days_burnout * mean_rate
  capacity = capacity_ratio * mean_rate
  
  #' Generate a sample case rate with a Poisson process
  create_swab_date = function(n, mean_rate) {
    swab_date = numeric(0)
    i = 1
    while (length(swab_date) < (n)) {
      # swab_date = c(swab_date, rep(i, rpois(1, mean_rate)))
      # Size parameter multiplier taken from Queue_Parameters.Rmd
      # swab_date = c(swab_date, rep(i, rnbinom(1, mu=mean_rate, size=mean_rate)))
      swab_date = c(swab_date, rep(i, rpois(1, lambda=mean_rate)))
      i = i + 1
    }
    swab_date = swab_date[seq_len(n)]
  }
  
  # Initialise population
  samples = tibble(
    swab_date = create_swab_date(n_burnin+n_samples+n_burnout, mean_rate),
    # notification_date = swab_date + pmin(rpois(n_burnin + n_samples + n_burnout, 1.5), 8),
    notification_date = swab_date + sample(x = test_turnaround_time$x,
                                           size = n_burnin + n_samples + n_burnout,
                                           replace = TRUE,
                                           prob = test_turnaround_time$p),
    notification_time = runif(n_burnin + n_samples + n_burnout),
    vaccinated = runif(n_burnin + n_samples + n_burnout) <= proportion_cases_vaccinated,
    interview_date = NA_integer_
  )
  
  # Run queue simulation
  # pb = txtProgressBar(min = 1, max = max(samples$notification_date), initial = 1)
  for (sim_day in seq_len(max(samples$notification_date))) {
    # setTxtProgressBar(pb, sim_day)
    samples = samples %>%
      mutate(eligible_for_interview = notification_date <= sim_day & # Notification has happened
               is.na(interview_date) & # Hasn't been interviewed already
               notification_date >= (sim_day-max_interview_delay) & # Notification hasn't expired
               !(notification_date==sim_day & notification_time < prop_time_delay)) %>%
      # Sort by priority function
      f_priority() %>%
      # Apply interview status
      mutate(interview_date = ifelse(
        row_number() <= capacity & eligible_for_interview,
        sim_day,         # Interview
        interview_date)) # No interview
  }
  
  # Remove burnin and burnout
  samples = samples[(n_burnin+1):(n_burnin+n_samples),] %>%
    mutate(
      samples_isol_swab = rpois(n(), 1),
      vaccinated = ifelse(vaccinated, "Pfizer", "None"),
      test_turnaround_time = notification_date - swab_date,
      time_to_interview = replace_na(interview_date - notification_date, Inf)
    )
}
```

Execute sim_tracing function convergence test

```{r, eval=T}
mean_rates = c(10, 20, 50, 100, 500, 1000)
capacities = c(0.2, 0.5, 0.8)
combinations = expand_grid(rate=mean_rates, capacity=capacities)
library(pbmcapply)
samples_list = pbmclapply(seq_len(nrow(combinations)),
                          function(i) {
                            rate = combinations$rate[i]
                            capacity = combinations$capacity[i]
                            sim_tracing(capacity_ratio = capacity,
                                        mean_rate = rate,
                                        f_priority = random_swab,
                                        days_samples = 1000) %>%
                              count(time_to_interview) %>%
                              mutate(p = n / sum(n),
                                     center = (n + 0.5*z^2) / (sum(n) + z^2),
                                     n = as.double(n),
                                     band = (z / (sum(n) + z^2)) * sqrt((n * (sum(n)-n)) / sum(n) + z^2/4),
                                     lower = center - band,
                                     upper = center + band,
                                     mean_rate = rate,
                                     capacity = capacity)
                          },
                          mc.cores = min(8, nrow(combinations)))
samples_list.2 = samples_list %>%
  bind_rows() %>%
  mutate(mean_rate = factor(mean_rate),
         capacity = paste("Capacity ratio =", capacity))

xmax = max(samples_list.2$time_to_interview[is.finite(samples_list.2$time_to_interview)])
samples_list.2$time_to_interview[samples_list.2$time_to_interview==Inf] = xmax+2

# cc = scales::seq_gradient_pal("navy", "tomato")(seq(0,1,length.out=length(mean_rates)))

samples_list.2 %>%
  ggplot(aes(x = time_to_interview, y = p, color = mean_rate, fill = mean_rate)) +
  geom_errorbar(aes(ymin = lower, ymax=upper), position=position_dodge(0.5)) +
  geom_point(position = position_dodge(0.5)) +
  facet_wrap(vars(capacity), ncol=1) +
  # scale_color_continuous(trans = "log", breaks=mean_rates, labels=scales::comma) +
  # scale_color_manual(values = cc) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = c(0:xmax, xmax+2), labels=c(0:xmax, "Missed")) +
  # theme_minimal() +
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank()) +
  labs(title = "Inspection of delay invariance",
       subtitle = "with respect to incoming case rate, using the random prioritisation strategy",
       x = "Overall interview delay (days)",
       y = "Proportion of cases",
       color = "Mean case\nrate",
       fill = "Mean case\nrate")
ggsave("plot1.svg", width=150, height=150, units="mm")
```

# Run experiments

```{r}
strategies = list(
  "Random" = random_swab,
  "Oldest swab" = oldest_swab,
  "Newest notification" = newest_notification,
  "Newest swab" = newest_swab,
  "Newest swab then unvaccinated" = new_swab_vaccine
)
capacity_ratios = list(
  0.2, 0.5, 0.8
)
mean_rate = 20
scenarios = cross(list(strategy = names(strategies),
                       capacity_ratio = capacity_ratios,
                       mean_rate = mean_rate))

strategy_samples = future_lapply(scenarios, function(s) {
  sim_tracing(mean_rate = s$mean_rate,
              capacity_ratio = s$capacity_ratio,
              days_samples = 1000,
              f_priority = strategies[[s$strategy]]) %>%
    mutate(time_to_interview = ifelse(time_to_interview < 0, Inf, time_to_interview))
},
future.seed=TRUE)

strategy_summary = lapply(strategy_samples, function(x) {
  x %>%
    count(time_to_interview) %>%
    mutate(p = n / sum(n),
           center = (n + 0.5*z^2) / (sum(n) + z^2),
           n = as.double(n),
           band = (z / (sum(n) + z^2)) * sqrt((n * (sum(n)-n)) / sum(n) + z^2/4),
           lower = center - band,
           upper = center + band,
           mean_rate = mean_rate)
})
# Assign scenario parameters back to summary
for (i in seq_along(strategy_summary)) {
  for (j in names(scenarios[[1]])) {
    strategy_summary[[i]][[j]] = scenarios[[i]][[j]]
  }
}

strategy_plot_data = bind_rows(strategy_summary) %>%
  mutate(strategy = strategy %>% str_replace("n u", "n\nu") %>% fct_inorder(),
         capacity_ratio_label = percent(capacity_ratio, suffix="% capacity"))
strategy_labels = strategy_plot_data %>%
  mutate(center = max(center),
         capacity_ratio = min(capacity_ratio)) %>%
  distinct(strategy, .keep_all=TRUE)
xmax = max(strategy_plot_data$time_to_interview[is.finite(strategy_plot_data$time_to_interview)])
strategy_plot_data$time_to_interview[is.infinite(strategy_plot_data$time_to_interview)] <- xmax + 2
ggplot(strategy_plot_data,
       aes(x = time_to_interview,
           y = center,
           fill = time_to_interview <= xmax)) +
  geom_col() +
  # geom_text(aes(x=0, y=center, label=strategy, color=NULL, fill=NULL, group=NULL),
  #           data=strategy_labels, hjust=0, vjust=1, color="grey20", size=3.5, nudge_x=-0.5, nudge_y=-0.05) +
  # geom_errorbar(aes(ymin = lower, ymax=upper), width=0.5) +
  facet_grid(rows = vars(strategy),
             cols = vars(capacity_ratio_label),
             switch = "y") +
  scale_x_continuous(breaks = c(0:xmax, xmax+2), labels=c(0:xmax, "Missed")) +
  scale_y_continuous(labels=percent_format()) +
  theme_minimal() +
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        # panel.background = element_rect(fill = NA),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none",
        strip.background.x = element_rect(fill="grey80", color=NA),
        strip.background.y = element_rect(fill="grey80", color=NA),
        strip.placement = "outside"
        # strip.text.y = element_blank()
  ) +
  labs(title = "Interview delay distributions for prioritisation strategies",
       x = "Interview delay (days from notification to interview)",
       y = "Proportion of cases",
       fill = "Interviewed\n(to show NAs)")
ggsave("plot2.svg", width=150, height=200, units="mm")
```

# TP reduction

Nick's method

```{r}
gi_meanlog = 1.375738
gi_sdlog = 0.5665299
p_active_detection = 0.95
passive_detection_given_symptoms = 0.5
passive_distribution = get_passive_distribution()

# (this will change as a function of vaccination coverage)
pr_symptoms = 0.3
p_passive_detection =  passive_detection_given_symptoms * pr_symptoms

strategy_samples = lapply(strategy_samples, function(df) {
  df = df %>%
    mutate(tracing_delay = samples_isol_swab +
             test_turnaround_time +
             time_to_interview)
})

strategy_isolation = future_lapply(strategy_samples, function(df) {
  library(distributional)
  time_to_isolation(10000,
                    gi_meanlog,
                    gi_sdlog,
                    p_active_detection,
                    p_passive_detection,
                    passive_distribution,
                    df) %>%
    as_tibble()
}, future.seed=TRUE)

# isolation_summary = lapply(strategy_isolation, function(x) {sum(is.infinite(x$time_to_passive))})
# x = strategy_isolation[[1]]
# 
# tp_multiplier = unlist(calculate_tp_multiplier_queue(strategy_isolation[[1]])$tp_multiplier_individual)

tp_reduction = function(df, n=10000) {
  tp_multiplier_individual = unlist(calculate_tp_multiplier_queue(df)$tp_multiplier_individual)
  mean_bootstrap = replicate(
    n,
    mean(sample(1-tp_multiplier_individual, length(tp_multiplier_individual), replace=TRUE)))
  mean_bootstrap_quantiles = quantile(mean_bootstrap, c(0.025, 0.975))
  tibble(mean = 1-mean(tp_multiplier_individual),
         lower = mean_bootstrap_quantiles[[1]],
         upper = mean_bootstrap_quantiles[[2]]
  )
}

#' Mean estimator using bootstrapping
# tp_reduction = function(df, n=10000) {
#   tp_reduction = tp_reduction_individual(df$time_to_interview)
#   mean_bootstrap = replicate(
#     n,
#     mean(sample(tp_reduction, length(tp_reduction), replace=TRUE)))
#   mean_bootstrap_quantiles = quantile(mean_bootstrap, c(0.025, 0.975))
#   tibble(mean = mean(tp_reduction),
#          lower = mean_bootstrap_quantiles[[1]],
#          upper = mean_bootstrap_quantiles[[2]]
#   )
# }

strategy_tp = future_lapply(strategy_isolation, tp_reduction, future.seed=T)
# Assign scenario parameters back to summary
for (i in seq_along(strategy_summary)) {
  for (j in names(scenarios[[1]])) {
    strategy_tp[[i]][[j]] = scenarios[[i]][[j]]
  }
}
strategy_tp = bind_rows(strategy_tp) %>%
  mutate(capacity_ratio_label = percent(capacity_ratio, suffix="% capacity"),
         strategy = strategy %>% str_replace(" (?!.* )", "\n") %>% fct_inorder())

ggplot(strategy_tp, aes(x=strategy, y=mean, label=percent_format(accuracy=1)(mean))) +
  geom_col(fill = "steelblue") +
  # geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5) +
  geom_text(hjust=0.5, vjust = 1, color = "white", nudge_y = -0.005, size=3) +
  facet_grid(cols = vars(capacity_ratio_label)) +
  scale_y_continuous(labels = scales::percent_format()) +
  # coord_flip() +
  labs(x = NULL,
       y = "Reduction in transmission") +
  theme_minimal() +
  theme(strip.background.x = element_rect(fill="grey80", color=NA),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))
ggsave("plot3.svg", width=150, height=75, units="mm")
```

Kretzschmar method

Read in estimates for TP reduction by lab/interview delays.

```{r}
# Define TP reductions
kretzschmar_lookup = read.csv("data-public/testing_delay_kretzschmar_table_2_extended.csv") %>%
  pivot_longer(-Testing.delay,
               names_to = "contact_tracing_delay",
               values_to = "kretzschmar_tp_reduction") %>%
  mutate(contact_tracing_delay = contact_tracing_delay %>%
           str_remove_all("[^[0-9]]") %>%
           as.numeric() %>%
           replace_na(-2))
# vaccine_tp_reduction = tribble(
#   ~vaccine, ~vaccine_tp_reduction,
#   "AZ", 0.505,
#   "Pfizer", 0.505,
#   "None", 0
# )
```

Apply TP reductions to each case

```{r, eval=F}
kretzschmar_reduction_individual = function(df) {
  df %>%
    # Calculate TP reductions
    # left_join(vaccine_tp_reduction,
    #           by = c("vaccinated" = "vaccine")) %>%
    left_join(kretzschmar_lookup,
              by = c("test_turnaround_time" = "Testing.delay",
                     "time_to_interview" = "contact_tracing_delay")) %>%
    pull(kretzschmar_tp_reduction)
}

#' Mean estimator
kretzschmar_tp_reduction = function(df, n=10000) {
  tp_reduction = kretzschmar_reduction_individual(df)
  tp_reduction = tp_reduction[!is.na(tp_reduction)] # Ignore any that go out of the bounds explored by Kretzschmar. Count is negligible (e.g. 1/20000) - alternatively could ensure RNGs do not go out of range.
  mean_bootstrap = replicate(
    n,
    mean(sample(tp_reduction, length(tp_reduction), replace=TRUE)))
  mean_bootstrap_quantiles = quantile(mean_bootstrap, c(0.025, 0.975), na.rm=TRUE)
  tibble(mean = mean(tp_reduction),
         lower = mean_bootstrap_quantiles[[1]],
         upper = mean_bootstrap_quantiles[[2]])
}
library(future.apply)
plan(multisession, workers = 4)
kretzschmar_strategy_tp = future_lapply(strategy_samples, kretzschmar_tp_reduction) %>%
  bind_rows(.id = "name")

# ggplot(kretzschmar_strategy_tp, aes(x=name, y=mean, label=round(mean, 3))) +
#   geom_col(fill = "steelblue") +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5) +
#   geom_text(vjust = 1, color = "white", nudge_y = -0.005) +
#   coord_cartesian(ylim = c(min(kretzschmar_strategy_tp$lower) - 0.05, NA)) +
#   labs(title = "Transmission potential reduction",
#        subtitle = "Kretzschmar method",
#        x = "Interview queue strategy",
#        y = "Reduction factor")
# ggsave("plot4.svg", width=75, height=75, units="mm")

bind_rows(Golding = strategy_tp,
          Kretzschmar = kretzschmar_strategy_tp,
          .id="method") %>%
  mutate(method = paste(method, "method")) %>%
  ggplot(aes(x=name, y=mean, label=round(mean, 3))) +
  geom_col(fill = "steelblue") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5) +
  geom_text(vjust = 1, color = "white", nudge_y = -0.005) +
  coord_cartesian(ylim = c(min(kretzschmar_strategy_tp$lower) - 0.025, NA)) +
  facet_wrap(vars(method), ncol=1, scales="free_y") +
  labs(title = "Transmission potential reduction with an under-resourced contact tracing queue",
       x = "Interview queue priority",
       y = "Reduction factor") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank())
ggsave("plot5.svg", width=150, height=75, units="mm")
```

