---
title: "Queue TP Reduction"
author: "Logan Wu"
date: "`r format(lubridate::today(), '%B %e, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(patchwork)
library(readxl)
library(lubridate)
library(distributional)
library(future.apply)
library(truncdist)
library(scales)
library(janitor)
library(patchwork)
plan(multisession)
z = 1.96

for(x in list.files("R", full.names=T)) {source(x)}
```

This RMarkdown illustrates the effectiveness of queueing strategy on transmission potential (TP) reduction in a contact tracing context.

```{r}
data_dir = "data"
VIC = read_cases_vic(file.path(data_dir, "Linelist_Cases_20210917.xlsx")) %>%
  rename(swab_date = earliest_detected,
         notification_date = earliest_confirmed_or_probable,
         interview_date = interviewed_date)
NSW = read_cases_nsw(file.path(data_dir, "CASES_FROM_20200701_0000_TO_20210913_1115.xlsx")) %>%
  rename(swab_date = earliest_detected,
         notification_date = earliest_confirmed_or_probable,
         interview_date = interviewed_date)
```

# Incoming cases

Calculate parameters for NB(r, mu)

```{r}
window_width = 14
NSW_optimal = NSW %>%
  count(notification_date) %>%
  mutate(moving_average = zoo::rollapply(n, window_width, mean, align="center", fill=NA)) %>%
  filter(notification_date >= as.Date("2020-07-01"),
         notification_date <= as.Date("2021-01-31")) %>%
  drop_na(moving_average)

likelihood_fn = function(r) {
  ll = sapply(seq_len(nrow(NSW_optimal)), function(i) {
    mean = NSW_optimal$moving_average[i]
    actual = NSW_optimal$n[i]
    dnbinom(actual, size=r, mu=mean, log=T)
  })
  sum(ll)
}

max_likelihood = optim(1000, likelihood_fn, lower=1e-6, method="L-BFGS-B")

size = max_likelihood$par
print(paste("Size:", size))
NSW_optimal_banded = NSW_optimal
alpha1 = 0.05
alpha2 = 0.5
NSW_optimal_banded$lower1 = sapply(NSW_optimal_banded$moving_average, function(x) {qnbinom(alpha1/2, size=size, mu=x)})
NSW_optimal_banded$upper1 = sapply(NSW_optimal_banded$moving_average, function(x) {qnbinom(1-alpha1/2, size=size, mu=x)})
NSW_optimal_banded$lower2 = sapply(NSW_optimal_banded$moving_average, function(x) {qnbinom(alpha2/2, size=size, mu=x)})
NSW_optimal_banded$upper2 = sapply(NSW_optimal_banded$moving_average, function(x) {qnbinom(1-alpha2/2, size=size, mu=x)})
NSW_optimal_banded$lower3 = sapply(NSW_optimal_banded$moving_average, function(x) {qpois(alpha1/2, lambda=x)})
NSW_optimal_banded$upper3 = sapply(NSW_optimal_banded$moving_average, function(x) {qpois(1-alpha1/2, lambda=x)})
NSW_optimal_banded$lower4 = sapply(NSW_optimal_banded$moving_average, function(x) {qpois(alpha2/2, lambda=x)})
NSW_optimal_banded$upper4 = sapply(NSW_optimal_banded$moving_average, function(x) {qpois(1-alpha2/2, lambda=x)})
# mutate(lower = qnbinom(0.25, size=size, mu=moving_average),
#        upper = qnbinom(0.75, size=size, mu=moving_average))

p1 = ggplot(NSW_optimal_banded, aes(x=notification_date, y=moving_average)) +
  geom_ribbon(aes(ymin=lower1, ymax=upper1, fill="95%"), alpha=0.3, fill="steelblue") +
  geom_ribbon(aes(ymin=lower2, ymax=upper2, fill="50%"), alpha=0.9, fill="steelblue") +
  geom_line() +
  geom_point(aes(y=n), alpha=0.5, size=1) +
  scale_x_date(date_labels = "%b %Y") +
  labs(subtitle = paste0("X ~ Neg. Binomial(μ(t), r=", round(size), ")"),
       x = "Notification date", y = "Cases")
p2 = ggplot(NSW_optimal_banded, aes(x=notification_date, y=moving_average)) +
  geom_ribbon(aes(ymin=lower3, ymax=upper3, fill="95%"), alpha=0.3, fill="tomato") +
  geom_ribbon(aes(ymin=lower4, ymax=upper4, fill="50%"), alpha=0.9, fill="tomato") +
  geom_line() +
  geom_point(aes(y=n), alpha=0.5, size=1) +
  scale_x_date(date_labels = "%b %Y") +
  labs(subtitle = paste("X ~ Poisson(μ(t))"),
       x = "Notification date", y = "Cases")
p1 + p2 +
  plot_layout(guides="collect") +
  plot_annotation(title="Maximum-Likelihood Fit for New South Wales",
                  subtitle=paste0("μ(t) is a ", window_width, "-day moving average"))

ggsave("plot4.svg", width=150, height=75, units="mm")
```

# Investigate notification delays

```{r}
linelist = bind_rows(VIC, NSW, .id="State") %>%
  mutate(notification_delay = notification_date - swab_date,
         State = ifelse(State=="1", "VIC", "NSW")) %>%
  filter(notification_date >= as.Date("2020-07-01"),
         notification_date <= as.Date("2021-09-30"))
```

Delays over time

```{r}
max_delay = 8
p1 = linelist %>%
  arrange(notification_date) %>%
  mutate(yearmonth = format(notification_date, "%b %Y"),
         yearmonth = fct_inorder(yearmonth)) %>%
  count(yearmonth, notification_delay, State) %>%
  filter(notification_delay >= 0,
         notification_delay < max_delay) %>%
  group_by(yearmonth) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = notification_delay, y = p, fill = State)) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  facet_wrap(~yearmonth) +
  labs(subtitle = "Notification delays for cases by month",
       x = NULL, y = "Proportion") +
  theme(legend.position = "none")

p2 = linelist %>%
  count(notification_delay, State) %>%
  filter(notification_delay >= 0,
         notification_delay < max_delay) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = notification_delay, y = p, fill = State)) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  labs(subtitle = "For all cases from July 2020 to September 2021",
       x = "Notification delay", y = NULL) +
  theme(legend.position = "bottom")

p1 / p2 + plot_layout(heights = c(3, 1))
ggsave("plot7.svg", width=150, height=150, units="mm")
```

Delays in the whole period

```{r}
delays = linelist %>%
  count(notification_delay) %>%
  mutate(notification_delay = as.integer(notification_delay),
         p = n / sum(n))
fit = delays %>%
  filter(notification_delay >= 0, notification_delay <= 8) %>%
  # group_by(State) %>%
  summarise(notification_delay = weighted.mean(notification_delay, n, na.rm=T),
            weight = sum(n))
mean_notification_delay_from_data = weighted.mean(fit$notification_delay, fit$weight)
fit_distribution = tibble(notification_delay = 0:8,
                          p = dpois(notification_delay, mean_notification_delay_from_data))

ggplot(delays, aes(x = notification_delay, y = p)) +
  geom_col() +
  geom_line(data=fit_distribution) +
  # geom_vline(data=fit, aes(xintercept=notification_delay)) +
  # facet_wrap(vars(State), scales = "free_y") +
  coord_cartesian(xlim = c(0, 5))
```

Write delays to a list to be read by the queue TP reduction RMD

```{r}
delays %>%
  filter(notification_delay >= 0,
         notification_delay < 14) %>%
  write_csv("notification_delays.csv")
```

# Investigate notification clustering

How many cases are reported per day

```{r}
notified = linelist %>%
  count(State, notification_date)

ggplot(notified, aes(x = notification_date, y = n, color = State)) +
  geom_line()
```
